---
title: "Activity"
author: "Setareh"
date: "10/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data


```{r echo=TRUE }

if (!file.exists('activity.csv')) {
  unzip(zipfile = "activity.zip")
}

activityData <- read.csv(file="activity.csv", header=TRUE)
```

## Including Plots


```{r , echo=TRUE}
##total number of steps per day

totalSteps <- aggregate(steps ~ date, activityData, FUN=sum)

### Make a histogram 
hist(totalSteps$steps,
     main = "Total Steps per Day",
     xlab = "Number of Steps")

## Calculate and report the mean and median of total steps taken per day
meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
medSteps <- median(totalSteps$steps, na.rm = TRUE)
```
## The average daily activity pattern 
```{r echo=TRUE }
library(ggplot2)

meanStepsByInt <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = meanStepsByInt, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))


 
maxInt <- meanStepsByInt[which.max(meanStepsByInt$steps),]

```
### Which 5-minute interval across all days contain the maximum number of steps
```{r echo=TRUE }
library(ggplot2)

meanStepsByInt <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = meanStepsByInt, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))

```
### Which 5-minute interval across all days contain the maximum number of steps
 
```{r echo=TRUE }

maxInt <- meanStepsByInt[which.max(meanStepsByInt$steps),]

```
##Imputing Missing Values
### Calculate and report the total number of missing values in the dataset

```{r echo=TRUE }
missingVals <- is.na(activityData$steps)

```
###Devise a strategy for filling in all of the missing values

### Create a new dataset that is equal to the original dataset but with 
###the missing data filled in.
```{r echo=TRUE }
imp_activityData <- transform(activityData,
                              steps = ifelse(is.na(activityData$steps),
                                             meanStepsByInt$steps[match(activityData$interval, 
                                                                        meanStepsByInt$interval)],
                                             activityData$steps))


```
### Make a histogram of the total number of steps taken each day and
###and report the mean and median.
```{r echo=TRUE }
impStepsByInt <- aggregate(steps ~ date, imp_activityData, FUN=sum)
hist(impStepsByInt$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Number of Steps")

impMeanSteps <- mean(impStepsByInt$steps, na.rm = TRUE)
impMedSteps <- median(impStepsByInt$steps, na.rm = TRUE)
diffMean = impMeanSteps - meanSteps
diffMed = impMedSteps - medSteps
diffTotal = sum(impStepsByInt$steps) - sum(totalSteps$steps)


```
##Are there differences in activity patterns between weekdays and weekends?

### Create a new factor variable in the dataset with two levels - "weekend" and "weekday"
```{r echo=TRUE }
DayType <- function(date) {
  day <- weekdays(date)
  if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
    return ("weekeday")
  else if (day %in% c('Saturday', 'Sunday'))
    return ("weekend")
  else
    stop ("Invalid Date Format.")
}
imp_activityData$date <- as.Date(imp_activityData$date)
imp_activityData$day <- sapply(imp_activityData$date, FUN = DayType)

```
### Make a panel plot containnig a time-series plot of the 5-minute interval
###and the average number of steps taken across all weekdays or weekends
```{r echo=TRUE }
meanStepsByDay <- aggregate(steps ~ interval + day, imp_activityData, mean)
ggplot(data = meanStepsByDay, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(day ~ .) +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))


```

