---
title: "peer assignment RA"
author: "Setareh"
date: "11/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##loading Library

```{r , echo=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
library(plyr)

```


##Download the file, unzip and read the .csv file.

```{r , echo=TRUE }
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp, mode="wb")
unzip(temp, "activity.csv")
activity <- read.csv("activity.csv",header=T)
unlink(temp)
```

##Calculate the total number of steps taken per day?
###(For this part of the assignment, you can ignore the missing values in the dataset.)

```{r , echo=TRUE}
totalstepsperday <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
head(totalstepsperday)

```

##Make a histogram of the total number of steps taken each day. convert dates first

```{r , echo=TRUE}
## converting dates to Y-M-D format
activity$date <- as.Date(activity$date, "%Y-%m-%d")
## calculate steps as it relates to date using SUM (per day)
hist(totalstepsperday$steps, 
    main="Total Steps per Day", 
    xlab="Number of Steps per Day", 
    ylab = "Interval",
    col="orange",
    breaks=50)
```

##Calculate and report the mean and median total number of steps taken per day.

```{r , echo=TRUE}
## mean of total steps per day
msteps <- mean(totalstepsperday$steps)
msteps
```

```{r , echo=TRUE}
## median of total steps per day
medsteps <- median(totalstepsperday$steps)
medsteps
```

```{r , echo=TRUE}
## check work using summary
summary(totalstepsperday)
```

##What is the average daily activity pattern?
###Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r , echo=TRUE}
## five minute average using steps to interval - FUN = mean instead of sum
fivemin <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)
## line chart
plot(x = fivemin$interval, 
    y = fivemin$steps, 
    type = "l", 
    col = "orange",
    xlab = "5-minute Intervals",
    ylab = "Average Steps Taken ~ Days",
    main = "Average Daily Activity Pattern")
```

##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?


```{r , echo=TRUE}
maxsteps <- fivemin$interval[which.max(fivemin$steps)]
maxsteps
```
##Imputing missing values
##Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r , echo=TRUE}
activity2 <- activity
nas <- is.na(activity2$steps)
avg_interval <- tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify = TRUE)
activity2$steps[nas] <- avg_interval[as.character(activity2$interval[nas])]
names(activity2)
```

```{r , echo=TRUE}
## Check for no-NA
sum(is.na(activity2))
```

##Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r , echo=TRUE}
#Plotting
#Setting up the pannel for one row and two columns
par(mfrow=c(1,2))

## Similar analysis without NAs now
totalstepsperday2 <- aggregate(steps ~ date, data = activity2, FUN = sum, na.rm = TRUE)
head(totalstepsperday2)
```

```{r , echo=TRUE}
## Histogram without the NA values
hist(totalstepsperday2$steps, 
    main = "Total Steps per Day (no-NA)", 
    xlab = "Number of Steps per Day", 
    ylab = "Interval",
    col="green",
    breaks=50)
##Histogram with the orginal dataset
hist(totalstepsperday$steps, 
    main="Total Steps per Day (Original)", 
    xlab="Number of Steps per Day", 
    ylab = "Interval",
    col="orange",
    breaks=50)
```

```{r , echo=TRUE}
#Resetting the panel
par(mfrow=c(1,1))
## What is the impact of imputing data?
summary(totalstepsperday)
```

```{r , echo=TRUE}
summary(totalstepsperday2)

```

```{r , echo=TRUE}
## Data has three fields, and we will add a new one in the next step - 11
head(activity2)
```

```{r , echo=TRUE}
activity2<- activity2 %>%
        mutate(typeofday= ifelse(weekdays(activity2$date)=="Saturday" | weekdays(activity2$date)=="Sunday", "Weekend", "Weekday"))
head(activity2)
```

##Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r , echo=TRUE}
## Plot - Line chart
fivemin2<- aggregate(steps ~ interval, data = activity2, FUN = mean, na.rm = TRUE)
head(fivemin2)
```

```{r , echo=TRUE}
ggplot(activity2, aes(x =interval , y=steps, color=typeofday)) +
       geom_line() +
       labs(title = "Ave Daily Steps (type of day)", x = "Interval", y = "Total Number of Steps") +
       facet_wrap(~ typeofday, ncol = 1, nrow=2)
```

##reproduce the results (numbers, plots, etc.) in the report.


```{r , echo=TRUE}
message("All code was 'echo=TRUE'. Thanks for reading")

```

